---
title: "Lab 4"
author: "An Bui"
date: "1/31/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### 0. Attach packages. 
```{r set up, include = FALSE}
library(tidyverse)
library(Ecdat)
library(plm)
library(lmtest)
library(car)
```

#### 1. Non-linear least squares for logistic growth of bacteria
```{r bacteria data, include = FALSE}
#### read in file ####
cellgrowth <- read_csv("CellGrowth.csv")

#### visualize ####
cellgrowth_plot <- ggplot(cellgrowth, aes(x = Time, y = CellCount)) +
  geom_point()
cellgrowth_plot
# first 4 points look like exponential growth is dominant
```

Find estimates for logistic growth:
$$
N_t = \frac{A}{1 + Be^{-rt}}
$$
Estimate carrying capacity ~ 3700. If $N_0$ is 2000, what is B?
```{r find B}
BEstimate <- (3700 - 2000)/2000
BEstimate
```

Now, estimate $r$.
```{r estimate r}
#### wrangle ####
lag_phase <- cellgrowth %>% 
  filter(Time < 5)

#### plot ln cell growth vs time ####
lag_plot <- ggplot(lag_phase, aes(x = Time, y = log(CellCount))) +
  geom_point()
lag_plot
# linearly positive relationship between ln(cell count) and time
# slope of line is exponential rate constant r

#### find r ####
lag_phase_lm <- lm(log(CellCount) ~ Time, data = lag_phase)
lag_phase_lm
```
$r$ is about 0.1035.  

Remember that you have initial estimates for all variables:

* $r$ = 0.1035
* $K$ = 2700
* $N_0$ = 2000

Now, find A, B, and r using Gauss-Newton method for iterative convergence (nonlinear least squares).  

```{r use NLS function}
#### function ####
# use non-linear least squares (default is Gauss-Newton)
# enter variables even though R doesn't actually know what they are
cell_fit <- nls(CellCount ~ A/(1 + B*exp(-r*Time)),
                # give some starting value for each variable
                # which you already calculated up top
                start = list(A = 3700, B = 0.85, r = 0.1035),
                data = cellgrowth,
                # can see trace of iterations that R went through to get the convergence values
                trace = TRUE
                )
cell_fit
# gives trace pack of all the iterations
# values left of colon: square residual values
# at first level, initial estimates
# adjusts values at each level until it has reached convergence on parameter values


#### plot model on top of data ####
# set coefficients
A <- coef(cell_fit)[1]
B <- coef(cell_fit)[2]
r <- coef(cell_fit)[3]

# create new sequence of times
# sequence of range 0-20 of equal observations of length = 100 (100 total observations)
time_seq <- seq(0, 20, length = 100)

# plug into model
cell_pred <- A/(1 + B*exp(-r*time_seq))

# bind with time_seq data
pred_df <- data.frame(time_seq, cell_pred)

# create a graph with original data and model predictions
model_plot <- ggplot(cellgrowth, aes(x = Time, y = CellCount)) +
  geom_point(color = "blue", size = 3) +
  # add layer with model
  geom_line(data = pred_df, aes(x = time_seq, 
                                y = cell_pred),
            color = "orange",
            size = 1) 
model_plot
```

You can give model equation any parameters in any structure you want and give it starting values, and it will go through the iterative process to find convergent parameters.

#### 4. Panel Regression
Focus on:

* state (entity)
* year (time)

These two things together make this **panel data**. Also, focus on:  

* packpc (dependent variable)
* avgprs (predictor variable)

```{r cigarette data}
#### wrangle ####
cigs_panel <- Cigarette %>% 
  select(state, year, packpc, avgprs)

#### visualize ####
cigs_plot <- ggplot(cigs_panel, aes(x = avgprs, 
                                    y = packpc, 
                                    group = state)) +
  geom_point(aes(color = state)) +
  # add trendlines
  geom_smooth(method = "lm", aes(color = state), se = FALSE, size = 0.2)
cigs_plot
# generally, looks like as average price is affecting purchases in a similar way when comparing between states

#### what's happening over years? ####
cigs_year <- ggplot(cigs_panel, aes(x = year, 
                                    y = packpc, 
                                    group = state)) +
  geom_point(aes(color = state)) +
  geom_smooth(method = "lm", aes(color = state), se = FALSE, size = 0.2)
cigs_year
```

Do panel regression first only considering entity fixed effects, then add time fixed effects.

```{r entity fixed effects}
cigs_model <- plm(packpc ~ avgprs,
                  data = cigs_panel,
                  # give entity and time variables
                  index = c("state", "year"),
                  # do an entity fixed effect model
                  model = "within")
cigs_model
# coefficient is -0.32545

# robust standard error
coeftest(cigs_model, vcov. = vcovHC(cigs_model, type = "HC1"))
# gives statistical information for the robust standard error for estimates
# more conservative estimate of this relationship associated with standard error
```

Now, add time fixed effects.
```{r time fixed effects}
et_cigs_model <- plm(packpc ~ avgprs,
                     data = cigs_panel,
                     index = c("state", "year"),
                     # entity fixed effect
                     model = "within",
                     # time fixed effect
                     effect = "twoways")
et_cigs_model
# coefficient is -0.38872

#### has including a time fixed effect actually improved my model? ####
# use pFtest on both models, give more complete model first
pFtest(et_cigs_model, cigs_model)
# null hypothesis: model is not improved by adding time fixed effects
# reject null hypothesis because p-value

```















